Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

%%%%%%% Author sign off on this first submission, last revision
% Sou-Cheng Choi signed off on
% Yuhan Ding signed off on
% Fred J. Hickernell signed off on
% Jagadeeswaran Rathinavel signed off on
% Aleksei Sorokin signed off on 3/17/2023

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.

@@

\bibliographystyle{spmpsci}
\bibliography{FJH23,FJHown23,sc}

@@

The (Q)MC pipeline above can be decomposed into $4$ main components:
\begin{description}
    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.  May produce IID or low discrepancy sequences depending on the problem.
    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$.
    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.
    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
\end{description}
A variety of QMC materials and the their categorizations into the above framework are described in the following list.
\begin{itemize}
    %\item \cite{LEc2017a} provides a history of random number generation
    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
    \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
    \item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
    \item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades
    \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
    \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet}
    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
    \item Multi-level software \cite{GilesSoft,GilesQSoft}
    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others
    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
    \item CUBA \cite{CUBA}
\end{itemize}

>>>>

%%%%%%% Author sign off on this first submission, last revision
% Sou-Cheng Choi signed off on
% Yuhan Ding signed off on
% Fred J. Hickernell signed off on 3/17
% Jagadeeswaran Rathinavel signed off on
% Aleksei Sorokin signed off on 3/17/2023

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,KuoJoe08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.

@@

\bibliographystyle{spmpsci}
\bibliography{FJH23,FJHown23}

@@

The (Q)MC pipeline above can be decomposed into $4$ main components:
\begin{description}
    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.  May produce IID or low discrepancy sequences depending on the problem.
    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$.
    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.
    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
\end{description}
A variety of QMC materials and the their categorizations into the above framework are described in the following list.
\begin{itemize}
    %\item \cite{LEc2017a} provides a history of random number generation
    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
    \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
    \item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
    \item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades
    \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
    \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet}
    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
    \item Sobol' direction numbers \cite{JoeKuo03,KuoJoe08a,SobolDirection}
    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
    \item Multi-level software \cite{GilesSoft,GilesQSoft}
    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others
    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
    \item CUBA \cite{CUBA}
\end{itemize}
